{How do we copy and paste?\\The semantic drift of quotations in blogspace}

### 2014-12-15 brains-introduction.tex
finished correcting 4.2 with new formulas / legend of fig. 5 accordingly / addressed notes / other minor mods

The practical study of the transformation of public representations has emerged only recently.
For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago \citep{Claidiere07,claidiere2014darwinian}\marginnote{S: il y a beaucoup de meta-études philosophiques où ils reprennent des études empiriques, mais toujours peu de modèles. Peut-être que tu veux vérifer les deux que j'ai réjouté? il y a aussi éventuellement \citet{claidiere2009imitation} avec un peu de modèle, \citet{maccallum2012evolution} plus empirique, et \citet{claidiere2012effect} qui le commente}.
Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
%They exhibit several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint. One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations. However, these studies address transformations by focusing on the properties of the source of the quotation (\hbox{e.g.} news outlet {vs.} blog), or the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations. They essentially examine the effect of some properties of the quotation source (\hbox{e.g.} news outlet {vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus). Some diffusion-transformation models have been proposed, yet the very cognitive features which may determine or, at least, influence these transformations, are overlooked; which may appear to be relatively unsatisfying from a cognitive viewpoint.

@@

Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading} for an example of such models).
Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin
alphabet. The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, i.e. the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
A third psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
\citet{nelson2013activation}, finally, show the importance of clustering coefficient in a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free assocation).

>>>>

The practical study of the transformation of public representations has emerged only recently.
For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citet{Claidiere07,claidiere2014darwinian} as well as a hybrid empirical-theoretical protocol in \citet{maccallum2012evolution}) \seb{il y a beaucoup de meta-études philosophiques où ils reprennent des études empiriques, mais toujours peu de modèles. Peut-être que tu veux vérifer les deux que j'ai réjouté? il y a aussi éventuellement \citet{claidiere2009imitation} avec un peu de modèle, \citet{maccallum2012evolution} plus empirique, et \citet{claidiere2012effect} qui le commente}\cam{Bien vu MacCallum! Mais attends, quelles sont les \emph{2} références ajoutées? je vois Claidière 2014, en sus du précédent Claidière 2007; mais les autres sont à choisir dans la parenthèse... en tout cas, je penche pour citer Claidière 2007, 2014 et en sandwich MacCallum, avec une précision, cf. phrase précédente. If it suits you, you may kill these commentaries altogether!}
Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
%They exhibit several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint. One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations. However, these studies address transformations by focusing on the properties of the source of the quotation (\hbox{e.g.} news outlet {vs.} blog), or the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations. They essentially examine the effect of some properties of the quotation source (\hbox{e.g.} news outlet {vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus). Some diffusion-transformation models have been proposed, yet the very cognitive features which may determine or, at least, influence these transformations, are overlooked; which may appear to be relatively unsatisfying from a cognitive viewpoint.

@@

Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models (see for instance \citet{collins1975spreading}).
Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that word properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin
alphabet. The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, i.e. the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
A third psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
\citet{nelson2013activation}, finally, show the importance of clustering coefficient in a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free assocation).

### 2014-12-15 brains-protocol.tex
Note error in figures
finished correcting 4.2 with new formulas / legend of fig. 5 accordingly / addressed notes / other minor mods

Number of phonemes and number of syllables naturally exhibit a strong linear correlation ($0.8$).
The analysis showed a better prediction effect of number of phonemes over number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to focus the presented results on the former only.

Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see in Figure~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.

@@

@@

\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
    Each feature exhibits a specific and significant pattern favouring either high- or low-valued words for substitution.}
    \label{fig:feature-susceptibilities}
\end{figure*}

@@

Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival) - \phi(\wstart)$.
Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
$$\Delta_{\phi}(f) = \left< \phi(\warrival) - \phi(\wstart) \right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore) and $\phi(\wstart)=f$,  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
$$\Delta_{\phi}^0 (f) = \left<\phi\right>-f$$

@@

\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
    \caption{\textbf{Feature variation upon substitution:} average feature of the appearing word minus $\mathcal{H}_0$ vs. average feature of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
    The overall position of the curve with respect to $y = 0$ indicates the direction of the cognitive bias.
    The fact that all the curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.}
    \label{fig:feature-variations}
\end{figure*}

>>>>

The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.

Age of acquisition is a key variable which appears as a usual suspect in psycholinguistic studies.
Despite it being usually difficult to disentangle from many of the other variables, it is known to have independent effects, which is consistent with what we see on Figure~\ref{fig:feature-corrs-initial}: age of acquisition has a limited correlation to the other variables (absolute value not above $0.39$ if we exclude the number of syllables and the network properties), leading us to keep the variable in the rest of the analysis.

@@

\seb{Les figures des résultats sont en fait pour $\Delta t = 1$ jour, pas 2. Sauf contre-indication, je corrige les figures et le texte}
\cam{Oui très bien, n'hésite pas. Super pour les figures, je sais pas si l'inversion graphique q en haut q' en bas est facile à réaliser?}

@@

\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/feature_susceptibilities.png}
    \caption{\textbf{Substitution susceptibility:} average susceptibility to substitution vs. average feature value of a candidate word for substitution, with 95\% asymptotic confidence intervals.
    Each feature exhibits a specific and significant pattern favoring either high- or low-valued words for substitution.}
    \label{fig:feature-susceptibilities}
\end{figure*}

@@

Next, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$, i.e. $\phi(\warrival)$ as a function of $\phi(\wstart)$.
Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$:
$$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:
$$\nu_{\phi}^0 (f) = \left<\phi\right>$$

@@

\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
    \caption{\textbf{Feature variation upon substitution:} average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
    }
    \label{fig:feature-variations}
\end{figure*}
